\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2022-04-04 lin-vs-cen.tex
Pfffffffffft

Finally, \citet{kun.obrien.ea:polynomial} boldly conjecture that, for any graph $G$, $\chicen(G)\le 2\chilin(G)$.  However, before the current work, the only class of graphs for which centred chromatic number was known to be bounded by a linear function of linear chromatic number was bounded degree trees.  The current result adds further evidence for this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.
% 
% These notes are an attempt to improve the bound in \cref{kun_obrien_general} in the general case as well as for some special classes of graphs. First we point out that some new results can already reduce the exponent from $190$ to $39$ and eliminate the polylog factor.  (These improvements are already described by \citet[Theorem~1.9]{czerwinski.nadara.ea:improved}.)
% 
% First, \citet{chuzhoy.tan:towards} improved the exponent in the Excluded Grid Theorem:  There exists a constant $c$ such that any graph $G$ with $\tw(G)\ge k^{18}\log^c k$ has a $k\times k$ grid minor.  This already improves the bound from $k^{190}$ to $k^{90}$.
% Next, a recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).  Therefore, if $\td(G)\ge k^{19}$ then at least one of the following cases applies:
% \begin{enumerate}
%     \item $\tw(G)\ge k^{18}$: In this case the proof follows Steps~2 and 3 as before, but using the grid minor result of \citet{chuzhoy.tan:towards} to show that $\chilin(G)\in\Omega(k)$.
%     \item $G$ contains a tree $T$ of treedepth $k$.  The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
%     A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.  Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
% \end{enumerate}

% As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.  This, by itself, would reduce the exponent in \cref{kun_obrien_general2} from $39$ to $20$.  We observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.  For example, if we consider only planar graphs, then it is well known that any planar graph $G$ of treewidth $k$ contains an $\Omega(k\times k)$ grid minor.  This already reduces the exponent in \cref{kun_obrien_general2} further (for planar graphs) to $3$ ($2$ if we can prove the result for grid minors).  If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.
%

@@

% For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.

% For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.  For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.  The subgraph $G[S]$ of $G$ \emph{induced} by $S$ is the graph with vertex set $V(G[S]):=S\cap V(G)$ and edge set $E(G[S]):=\{vw\in E(G):\{v,w\}\subseteq S\}$.

@@

% For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.  When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.  More precisely $\mathcal{P'}:=\mathcal{P}\cup\{\{v\}:v\in V(G)\setminus(\cup P)\}$.

% A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.  Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.  A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.  For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.  Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.  A $\le\!\!\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has at most $\ell$ internal vertices.

@@

% \begin{lem}\label{normal_hall}
%   Let $G$ be a bipartite graph with parts $X$ and $Y$ such that, for any $A\subseteq X$, $|N_G(A)|\ge |A|+1$.  Then $G$ contains a subgraph $M$ (a matching) such that $\deg_M(x)=1$ for each $x\in X$ and $\deg_M(y)\le 1$ for each $y\in Y$.
% \end{lem}

@@

We will also make use of the following version of the Lovász Local Lemma:

@@

If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph.  Thus, our goal is to show that that $\chilin(G)\in \Omega(k)$ for any $k\times k$ pseudogrid.  We do this by showing that, any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$. We will do this in several steps (see \cref{outline}):

@@

\begin{lem}
  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.
  Then there exists $S\subseteq V(G)$ such that
  \begin{compactenum}[(i)]
    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|= 2$; and
    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
  \end{compactenum}
\end{lem}

@@

  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$, for each $i\in\{1,\ldots,t\}$.  The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.  Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears before any elements of $X_{\nu_i}$ in our random permutation of $Y$'' then the events $U_1,\ldots,U_k$ are mutually independent.  Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.  For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.  Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/25)$ random variable.  Therefore,
  \begin{align*}
    \Pr\left(\deg_H(\alpha)< 100r^2\right)
    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
    & \le \Pr(\binomial(t,1/25)\le 100r^2) \\
    & \le (2t)^{100r^2}(1/25)^{r^2}(1-1/25)^{t-100r^2} \\
    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
    & = \exp(ar^2 - bt)
  \end{align*}
  for fixed positive $a,b>0$.  Recall that we are working in the regime in which $|\Phi_\alpha|\ge d\ge Cr^2$ and $t\in\Omega(d/r^2)$ so, for sufficiently large $C$,
  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-\Omega(C|\Phi_\alpha|/r^2))$.
  \todo[inline]{Work the preceding calculation out}
  We are now ready for an application of \cref{weighted_lovasz}.\todo{Continue here}  For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $\Phi_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.  The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.

  \begin{align*}
      \Pr(E_\alpha)
      & \le \exp(ar^2-bt) \\
      & \le \exp(ar^2-c|\Phi_\alpha|/r^2) \\
      & \le \exp(|\Phi_\alpha|(ar^2/|\Phi_\alpha|)-c/r^2) \\
      & \le \exp(|\Phi_\alpha|(1-c/r^2) & \text{for $d>r^2/a$} \enspace .
  \end{align*}

  For each $\alpha\in X$, define $w(E_\alpha)=c/2r^2$ and observe that
  \begin{align*}
    p(1-p)^{|\Gamma_\alpha|}
    & \ge  p(1-p)^{?|\Phi_\alpha|} = \exp(\log c/2r^2 + ?|\Phi_\alpha|\log(1-c/2r^2))
  \end{align*}
  \todo[inline]{Sort out these calculations}

@@

  Now it is straightforward to verify that $Q:=Q_1\cup Q_2$ satisifies (\ref{hits_both}).  Furthermore, for each $\mu\in Q_1$, $B_r(\mu)$ contains no object in $Q_1\setminus\{\mu\}$ and contains at most one element of $Q_2$.  Similarly, for each $\mu\in Q_2$, $B_r(\mu)$ contains no element of $Q_2\setminus\{\mu\}$ and contains at most one element of $Q_1$.  Therefore $Q$ satisfies (\ref{spread_out}).
\end{proof}

\newpage

% \hrule
%
% \subsection{Subdivided Grids}
%
% Refer to \cref{grid_figure}.  For positive integers $a$ and $b$, the \emph{$a\times b$ grid} $G_{a\times b}$ is the graph with vertex set $V(G_{a\times b}):=\{1,\ldots,a\}\times\{1,\ldots,b\}$ and that contain an edge with endpoints $(v,w)$ and $(x,y)$ if and only if $|v-x|+|w-y|=1$.  For each $i\in\{1,\ldots,a\}$, the \emph{$i$th column} of $G_{a\times b}$ is the vertex set $\{(i,1),\ldots,(i,b)\}$ and, for each $j\in\{1,\ldots,b\}$, the $j$th row is the vertex set $\{(1,j),\ldots,(a,j)\}$.  We define the \emph{boundary} of $G_{a\times b}$ as $\partial G_{a\times b}:=\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.  We say that a path $P:=v_0,v_1,\ldots,v_\ell$ in $G_{a\times b}$ \emph{goes straight} at a vertex $v_i$, $i\in\{1,\ldots,k-1\}$ if $v_{i-1}$, $v_i$, and $v_{i+1}$ are contained in a single row or column.  Otherwise, we say that $P$ \emph{bends} at $v_i$.
%
% \begin{figure}
%   \begin{centre}
%     \begin{tabular}{c@{\hspace{2cm}}c}
%       \includegraphics{figs/grid-1} &
%       \includegraphics{figs/grid-2}
%     \end{tabular}
%   \end{centre}
%   \caption{The grid $G_{10}$ and its $1$-subdivision $G^{\onesub}_{10}$.}\
%   \label{grid_figure}
% \end{figure}
%
% % An $a'\times b'$ \emph{subgrid} of $G_{a\times b}$ is an induced graph of the form $G_{a\times b}[\{(v,w): v\in\{i,\ldots,i+a'-1\},\, w\in\{j,\ldots,j+b'-1\}$ for some $i\in\{1,\ldots,a-a'+1\}$ and $j\in\{1,\ldots,b-b'+1\}$.   The \emph{vertex boundary} of $G_{a\times b}$ is $\{1,a\}\times\{1,\ldots,b\}\cup \{1,\ldots,a\}\times\{1,b\}$.  The \emph{edge boundary} of $G_{a,b}$ consists of all edges incident to at least one vertex in the vertex boundary of $G_{a,b}$. The \emph{boundary} of $G_{a\times b}$ is the union of the edge and vertex boundaries.
%
%
% % Let $G^{\onesub}_k$ denote the $1$-subdivision of $G_k$.
%
% A \emph{simple $k\times k$ pseudogrid} is any graph that can be obtained from $G^{\onesub}_k$ by replacing some degree-$4$ vertices of $G_k$ with small binary trees.  Specifically, we may replace a degree-$4$ vertex with a tree whose leaves are the vertices in $N_{G^{\onesub}_k}(v)$ and whose two internal nodes $v_a$ and $v_b$ are each adjacent to two leaves and to each other.
%
% Let $G_0$ be a simple $k\times k$ pseudogrid and observe that there is a natural mapping $f:E(G)\to E(G_k)\cup V(G_k)$ which is fairly obvious except that any edge $v_av_b\in E(G_0)$ introduced when replacing a degree-$4$ vertex $v$ in $G_k$ has $f(v_av_b):= v$.
%
% \begin{lem}\label{one_path}
%    Let $S:=\bigcup_{i=1}^r S_i$ be a subset of $E(G_0)$ with the following properties:
%    \begin{compactenum}[(a)]
%      \item $|S_i|\le 2$ for each $i\in\{1,\ldots,r\}$; and
%      \item for each distinct $i,j\in\{1,\ldots,r\}$, $\dist_{G_k}(\bigcup_{e\in S_i}f(e), \bigcup_{e\in S_j}f(e))\ge C$.\todo{Choose the value of $C$}
%    \end{compactenum}
%    Then $G_0$ contains a path that contains each edge in $S$.
% \end{lem}
%
% \begin{proof}
%   To do.
% \end{proof}
%
%
% % For brevity, we use $G_{k}:=G_{k\times k}$ as a shorthand for the $k\times k$ grid.  We will work extensively with the $1$-subdivision $G^{\onesub}_k$ of $G_k$.
% %
% %
% % In this case, we define $\partial G^{\onesub}_k:= \partial G_k\cup N_{G^{\onesub}_k}[\partial G_k]$.\todo{Define open and closed neighbourhods $N_G(\cdot)$ and $N_G[\cdot]$.}
% %
% % \begin{lem}\label{seven_by_seven}
% %   Let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{7})\setminus \partial G^{\onesub}_7$ with $|I\cup L\cup Q|\le 3$ that have the following properties:
% %   \begin{compactenum}[(i)]\setcounter{enumi}{0}
% %     \item Each vertex is $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
% %     \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
% %     \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
% %   \end{compactenum}
% %   Then for any pair of boundary vertices $s$ and $t$, $G_{10}$ contains a path $P$ that
% %   \begin{compactenum}[(a)]
% %     \item begins at $s$ and ends at $t$;
% %     \item contains every vertex in $I\cup L\cup Q$;
% %     \item goes straight through each vertex in $I$; and
% %     \item bends at each vertex in $L$.
% %     % \item contains no boundary vertices or edges except $s$ and $t$.
% %   \end{compactenum}
% % \end{lem}
% %
% % \begin{proof}
% %   Case analysis(?).
% % \end{proof}
% %
% % It will be helpful to take some natural definitions for $G_k$ and extend them to $G^{\onesub}_k$.  To do this, we associate each vertex $v$ of $G^{\onesub}_k$ with a vertex $v^{\leftup}$ of $G_k$. For each vertex $v\in V(G_k)$, we define $v^{\leftup}:=v$. For each subdivision vertex $s$ in $G^{\onesub}_k$ with neighbours $v$ and $w$, we define $s^{\leftup}$ to be the lexicographically smaller of $v$ or $w$. In other words, $s^{\leftup}$ is the vertex of $G_k$ that is directly below or the left of $s$.
% %
% % For two vertices $v_1,v_2\in V(G^{\onesub}_k)$, with $v^{\leftup}_1:=(x_1,y_1)$ and $v^{\leftup}_2:=(x_2,y_2)$ we define the $L_\infty$-distance between $v_1$ and $v_2$ as $\|v_1v_2\|_\infty = \max\{|x_1-x_2|, |y_1-y_2|\}$. Refer to \cref{ball_figure}.  For each $v\in V(G^{\onesub}_k)$, the \emph{($L_\infty$) ball} of radius $\beta$ centred at $v$ is $B_{\beta}(v):=\{w\in V(G^{\onesub}_k):\|vw\|_\infty \le \beta\}$.  Because it comes up in several places, it is worth noting that $|B_{\beta}(v)|\le 4(\beta+1)^2$, and this bound is tight for each $v\in\{\beta+1,\ldots,\beta-r-1\}\times\{\beta+1,\ldots,b-\beta-1\}$.
% %
% % \begin{figure}
% %   \begin{centre}
% %     \includegraphics{figs/grid-3}
% %   \end{centre}
% %   \caption{The ball $B_2(v)=B_2(s)$ in $G^{\onesub}_{10}$.}
% %   \label{ball_figure}
% % \end{figure}
% %
% %
% % \begin{lem}\label{four_points}
% %   Let $P$ be a set of at most $3$ vertices of $G^{\onesub}_k$.  Then there exists a set $S$ of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.\todo{Define the interior of a ball.}
% % \end{lem}
% %
% % \begin{proof}
% %   Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.  Let $X:=\bigcup_{i=1}^t \{x_i\bmod 7,(x_{i+1})\bmod 7\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 7,(y_i+1)\bmod 7\}$.  Then $|X|,|Y|\le 2t\le 6$.  In particular, there exists some $x\in\{0,\ldots,6\}\setminus X$ and some $y\in\{0,\ldots,6\}\setminus Y$. For each $i,j\in\Z$, let $c_{i,j}:=(7i+x+3,7j+y+3)$ and observe that the ball $B_{3}(c_{i,j})$ does not have any point of $P$ on its boundary.  Furthermore, the set of balls $S:=\{B_{3}(c_{i,j}):i,j\in\Z\}$ is a partition of $V(G^{\onesub}_{k})$.  Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
% % \end{proof}
% %
% % \begin{lem}
% %   Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.  Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
% % \end{lem}
% %
% % \begin{proof}
% %   Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$. Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $18$.  Therefore each component of $H$ contains at most $3$ vertices.
% %
% %   For each component $C$ of $H$, apply \cref{four_points} to $V(C)$ and keep only those balls in $S$ that contain at least one vertex of $V(C)$.  This produces a set $S_C$ of disjoint balls, each of radius $3$, that each contain some vertex in $V(C)$ in their interior.  Any two balls $B_{C_1}$ and $B_{C_2}$ obtained from different components of $H$ are disjoint because the first ball contains a vertex $p_1\in P$ and the second ball contains a vertex $p_2\in P$ such that $\|p_1p_2\|\infty > 12$, but the balls $s_1$ and $s_2$ each have radius $3$.
% % \end{proof}
% %
% %
% % \begin{lem}\label{one_path_enchilada}
% %   Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
% %   \begin{compactenum}[(1)]\setcounter{enumi}{-1}
% %     \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
% %     \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
% %     \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
% %     \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
% %   \end{compactenum}
% %   Then $G^{\onesub}_k$ contains a path $P$ that
% %   \begin{compactenum}[(a)]
% %     \item contains all the vertices in $I\cup L\cup Q$;
% %     \item goes straight through each vertex in $I$; and
% %     \item turns at each vertex in $L$.
% %   \end{compactenum}
% % \end{lem}
% %
% % \begin{proof}
% %   Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$. Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.  Each such ball is intersected by exactly one row of the snakelike path.  Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in each ball that it intersects.
% % \end{proof}
% %
% %
% % \subsection{Approximate Voronoi Diagrams in $G^{\onesub}_k$}
% %
% %
% % % \todo[inline]{This lemma is not true. Pick a centre $x$ and place a point in each of the quadrants around $x$ where the $i$th point is at distance $10\beta - (ir)$ from $x$.  Then $x$ is incident on all four Voronoi regions.  We really have to use the fact that $\beta$ (or rather, $\beta+ar$) is also an upper bound on the distance we actually care about.}
% % \begin{lem}\label{voronoi_mapping}
% %   Let $\beta, k\in\N$, let $P\subseteq V(G^{\onesub}_k)$ be such that $\min\{\|pq\|_\infty:p,q\in\binom{P}{2}\} \ge \beta$, and define  $Q:=\bigcup_{p\in P} B_\beta(p)$. Then there exists a function $\vor:Q\to P$ such that
% %   \begin{compactenum}[(i)]
% %     \item \label{real_close} For each $p\in P$ and each $v\in B_{\beta/8-1}(p)$, $\vor(v)=p$.
% %     \item \label{never_far} For each $v\in Q$, $\|v\vor(p)\|_\infty \le 7\beta/4$.
% %     \item \label{three_in_ball} For any ball $B$ of radius $\beta/12$, $|\{\vor(v):v\in B\cap Q\}|\le 3$.
% %     \item \label{twocolour_neighbourhood} For any $v\in Q$, $|\{\vor(v)\}\cup\{\vor(w):w\in Q\cap N_{G^{\onesub}_k}(v)\}|\le 2$.
% %   \end{compactenum}
% % \end{lem}
% %
% % \begin{proof}
% %   We first define a function $\vor':Q\to P$ that satisfies (\ref{real_close})--(\ref{three_in_ball}) and then show how local modifications can be used to obtain a function $\vor$ that also satisfies (\ref{twocolour_neighbourhood}).
% %
% %   Arbitrarily order the vertices of $P$ as $p_1,\ldots,p_{|P|}$.  For each $v\in Q$, let $d(v):=\min\{\|vp\|_\infty:p\in P\}$, let $S(v):=\{p\in P:\|vp\|_\infty \le d(v)+3\beta/4\}$, let $i(v):=\min\{i:p_i\in S(v)\}$, and let $\vor'(v)=p_{i(v)}$.  In words, $d(v)$ represents the distance from $v$ to its nearest neighbour in $P$, $S(v)$ represents the set of approximate nearest neighbours of $v$ in $P$ (up to an additive error of $\beta/12$), and $\vor'(v)$ is the approximate nearest neighbour of $v$ of smallest index.
% %
% %   To see that $\vor'$ satisfies (\ref{real_close}), observe that, for each distinct $p,q\in P$ and $v\in B_{\beta/8-1}(p)$, the triangle inequality implies that $\|vq\|_\infty \ge \|pq\|_\infty-\|vp\|_\infty \ge \beta-\beta/8+1 > 7\beta/8$.  On the other hand $d(v)\le \beta/8$, so $d(v)+3\beta/4\le 7\beta/8$.  Therefore, $q\not\in S(v)$ for any $q\in P\setminus\{p\}$, so $\vor'(v)=p$.
% %
% %   That $\vor'$ satsifies (\ref{never_far}) follows immediately from the fact that each $v\in Q$, so $d(v)\le \beta$, and therefore $\|pv\|_\infty\le 7\beta/4$ for each $p\in S(v)$.
% %
% %   We now argue that $\vor'$ satisfies (\ref{three_in_ball}).  Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.  Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.  By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
% %   Since $v_1\in Q$, $d(v)\le\beta$ so
% %   \[
% %     \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
% %   \]
% %   Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
% %   \[
% %     \|p_{i_2}v_2\|_\infty
% %      < \|p_{i_1}v_2\|_\infty - 3\beta/4

>>>>

Finally, \citet{kun.obrien.ea:polynomial} boldly conjecture that, for any graph $G$, $\chicen(G)\le 2\chilin(G)$.  However, before the current work, the only class of graphs for which centred chromatic number was known to be bounded by a linear function of linear chromatic number was bounded degree trees.  The current result adds further evidence for this conjecture by establishing that, when $G$ is a $k\times k$ grid (or even a pseudogrid), then $\chicen(G)\in\Theta(\chilin(G))$.

@@

@@

@@

@@

We will also make use of the following weighted version of the Lovász Local Lemma:

@@

If some graph contains a $k\times k$ grid minor then it contains a $k\times k$ pseudogrid as a subgraph.  Therefore to prove \cref{kun_obrien_general2}, it suffices to prove:

\begin{lem}\label{pseudogrid_lower_bound}
  For any $k\times k$ pseudogrid $G$, $\chilin(G)\in\Omega(k)$.
\end{lem}

We do this by showing that for sufficiently small $\epsilon >0$,  any $\epsilon k$-colouring of any $k\times k$ pseudogrid $G$ contains an uncentred path $P$. We will do this in several steps (see \cref{outline}):

@@

\begin{lem}\label{doubled_colour_set}
  For every integer $r\ge 1$ there exists an integer $d\in O(r^4)$ such that the following is true, for every integer $k\ge 1$.
  Let $G$ be a $k\times k$ pseudogrid with grid-partition $\mathcal{P}:=\{V(P_\mu):\mu\in\VE(G_k)\}$ and let $\varphi$ be a vertex colouring of $G$ with the property that, for each $A\subseteq\varphi(V(G))$, $|\varphi_{\mathcal{P}}^{-1}(A)|\ge d|A|$.
  Then there exists $S\subseteq V(G)$ such that
  \begin{compactenum}[(i)]
    \item \label{hits_both} for each $\alpha\in\varphi(V(G))$, $|\{v\in S:\varphi(v)=\alpha\}|= 2$; and
    \item \label{spread_out} $|\tilde{B}_r(v)\cap S|\le 2$ for each $v\in S$.
  \end{compactenum}
\end{lem}

@@

  Let $\{\mu_1',\ldots,\mu_t'\}$ be the subset of $Y$ chosen by this process and let $\{\nu_1,\ldots,\nu_r\}\subseteq\Phi_\alpha$ be objects with $\nu_i\in\VE(G_r(\mu_i'))$, for each $i\in\{1,\ldots,t\}$.  The important observation now is that the sets $X_{\nu_1},\ldots,X_{\nu_t}$ are disjoint.  Therefore, if we let $U_i$ denote the event ``$\mu_i'$ apppears before any elements of $X_{\nu_i}$ in our random permutation of $Y$'' then the events $U_1,\ldots,U_k$ are mutually independent.  Indeed, each $U_i$ depends only on the relative order of $\{\mu_i'\}\cup X_{\nu_i}$ in the random permutation of $Y$.  For each $i\in\{1,\ldots,k\}$, $\Pr(U_i)\ge 1/25$.  Therefore, $\sum_{i=1}^k \mathbbm{1}_{U_i}$ dominates a $\binomial(k,1/25)$ random variable.  Therefore,
  \begin{align*}
    \Pr\left(\deg_H(\alpha)< 100r^2\right)
    & \le \Pr\left(\sum_{i=1}^t \mathbbm{1}_{U_i} < 100r^2\right) \\
    & \le \Pr(\binomial(t,1/25)\le 100r^2) \\
    & \le (2t)^{100r^2}(1/25)^{r^2}(1-1/25)^{t-100r^2} \\
    & \le \exp(100r^2\log(2t)+(t-100r^2)\log(1-1/?)) \\
    & \le \exp(-100r^2\log(2t)\log(1-1/?) + t\log(1-1/25)) \\
    & = \exp(ar^2 - bt)
  \end{align*}
  for fixed positive $a,b>0$.  Recall that we are working in the regime in  $t\in\Omega(|\Phi_\alpha|/r^2)$ and $|\Phi_\alpha|\in\Omega(d/r^2)$ so, for any fixed $C>0$, there exists $d\in\Omega(r^4)$ such that
  $\Pr(\deg_H(\alpha)< 100r^2)\le \exp(-C|\Phi_\alpha|/r^2)$.
  \todo[inline]{Work the preceding calculation out}
  We are now ready for an application of \cref{weighted_lovasz}.\todo{Continue here}  For each $\alpha\in X$, let $E_\alpha$ denote the event ``$\deg_H(\alpha)\le 100r^2$.''  For each $\alpha\in X$, define $\Gamma_\alpha$ to be the set that contains every event $E_\beta$ such that $\Phi_\alpha$ contains an element $\nu$ with $B_{10r}(\nu)\cap Z_\beta\neq\emptyset$.  The event $X_\alpha$ is mutually independent of $\{X_\beta:\beta\in X\setminus\{\alpha\}\setminus\Gamma_\alpha\}$ since, for any $E_\beta\not\in \Gamma_\alpha$ the sets $Y_\alpha:=\{\mu\in Y:B_{2r}(\mu)\cap \Phi_\alpha\neq\emptyset\}$ and $Y_\beta:=\{\mu\in Y:B_{2r}(\mu)\cap Z_\beta\neq\emptyset\}$ are disjoint and the event $E_\alpha$ is entirely determined by the relative order of $Y_\alpha$ in the permutation while $E_\beta$ is entirely determined by the relative order of $Y_\beta$.\todo{Explain better}   Finally, observe that $|\Gamma_\alpha|\le 25|\Phi_\alpha|$.

  \begin{align*}
      \Pr(E_\alpha) \le \exp(-|\Phi_\alpha|\cdot C/r^2) \enspace .
  \end{align*}

  For each $\alpha\in X$, define $w(E_\alpha)=C/2r^2$ and observe that
  \begin{align*}
    p(1-p)^{|\Gamma_\alpha|}
    & \ge  p(1-p)^{25|\Phi_\alpha|} = p\exp(|\Phi_\alpha|\cdot 25\log(1-c/2r^2))
  \end{align*}
  \todo[inline]{Sort out these calculations}

@@

  Now it is straightforward to verify that $Q:=Q_1\cup Q_2$ satisifies (\ref{hits_both}).  Furthermore, for each $\mu\in Q_1$, $B_r(\mu)$ contains no object in $Q_1\setminus\{\mu\}$ and contains at most one element of $Q_2$.  Similarly, for each $\nu\in Q_2$, $B_r(\nu)$ contains no element of $Q_2\setminus\{\nu\}$ and contains at most one element of $Q_1$.  Therefore $Q$ satisfies (\ref{spread_out}).
\end{proof}

We now have all the pieces needed to complete our lower bound on the linear chromatic number of pseudogrids:

\begin{proof}[Proof of \cref{pseudogrid_lower_bound}]
  Suppose for the sake of contradiction that $G$ has a linear colouring $\varphi$ using fewer than $\epsilon k$ colours for arbitrarily small $\epsilon >0$.  Then, by \cref{only_frequent}, $G$ contains a $k'\times k'$ pseudogrid $G'$ with $k'\ge k/2$ having a grid-decomposition $\mathcal{P}':=\{P_\mu':\mu\in \VE(G_{k'})\}$ with the property that, for any $A\subseteq\varphi(V(G'))$, $|\varphi^{-1}_{\mathcal{P'}}(A)|\ge d|A|$, for any $d\le 1/2\epsilon$.  In particular, some $\epsilon \in \Omega(1/r^4)$, \cref{doubled_colour_set} implies that $G'$ contains a set $S$ of $2|\varphi(V(G')|$ vertices containing two vertices of each colour in $\varphi(V(G'))$ such that $B_r(\mu)\cap S = \{\mu\}$ for each $\mu\in S$.  By By \cref{pick_up_everything}, $G'$ contains a path $P$ that contains every vertex in $S$.  Since $P$ is contained in $G'$ $\varphi(V(P))=\varphi(V(G'))$, so $P$ is has no center under $\varphi$, contradicting the assumption that $\varphi$ is a linear colouring of $G$.
\end{proof}
